---
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

# `r titles[1]` -- `r report_year[1]` Lake Summary

```{r condition_summary, echo=FALSE, message=FALSE, warning=FALSE}
trophic_df <- lake1 %>% 
    distinct(SAMPLE_ID, INFO_TYPE, Characteristic.Name, Result.Sample.Fraction, Result.Value, .keep_all = TRUE)

trophic_df$Result.Sample.Fraction <- replace_na(trophic_df$Result.Sample.Fraction, "None")

trophic_df <- trophic_df %>% 
  mutate(FULL_CHARACTERISTIC_NAME = paste(Characteristic.Name, INFO_TYPE, Result.Sample.Fraction, sep=" ")) %>%
  filter(FULL_CHARACTERISTIC_NAME %in% c("DEPTH, SECCHI DISK DEPTH SD None",
                                         "PHOSPHORUS OW T",
                                         "CHLOROPHYLL A OW T"
                                         ))

for(i in 1:nrow(trophic_df)){
  if(trophic_df$SAMPLE_DATE[i] >= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 07, 15, sep = "-")) & trophic_df$SAMPLE_DATE[i] <= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 09, 15, sep = "-"))){
    trophic_df$season[i] <- "productivity"
  } else if(trophic_df$SAMPLE_DATE[i] >= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 06, 1, sep = "-")) & trophic_df$SAMPLE_DATE[i] <= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 07, 14, sep = "-"))){
    trophic_df$season[i] <- "summer"
  } else if(trophic_df$SAMPLE_DATE[i] >= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 09, 15, sep = "-")) & trophic_df$SAMPLE_DATE[i] <= as.Date(paste(year(trophic_df$SAMPLE_DATE[i]), 10, 1, sep = "-"))){
    trophic_df$season[i] <- "summer"
  } else {
   trophic_df$season[i] <- "not"
  }
}

trophic_df$Result.Value <- as.numeric(trophic_df$Result.Value)

trophic_df <- trophic_df %>% 
  filter(season != "not") %>% 
  filter(!is.na(Result.Value)) %>% 
  group_by(FULL_CHARACTERISTIC_NAME, year, season) %>% 
  dplyr::summarize(avg_result_s = mean(Result.Value)) %>% 
  ungroup() %>% 
  group_by(FULL_CHARACTERISTIC_NAME) %>%
  dplyr::summarize(LongtermAvg = mean(avg_result_s), standard_dev = sd(avg_result_s))

secchi <- trophic_df$LongtermAvg[trophic_df$FULL_CHARACTERISTIC_NAME == "DEPTH, SECCHI DISK DEPTH SD None"]



if(is.na(secchi)){
  secchi_score <- "0"
  secchi_reason <- "unsampled"
} else if(secchi >5){
  secchi_score <- 3
  secchi_reason <- "high"
} else if(secchi <2){
  secchi_score <- 1
  secchi_reason <- "low"
} else{
  secchi_score <- 2
  secchi_reason <- "moderate"
}

tphos <- trophic_df$LongtermAvg[trophic_df$FULL_CHARACTERISTIC_NAME == "PHOSPHORUS OW T"]

if(is.na(tphos)){
  tphos_score <- "0"
  tphos_reason <- "unsampled"
} else if(tphos > 0.02){
  tphos_score <- 1
  tphos_reason <- "high"
} else if(tphos < 0.01){
  tphos_score <- 3
  tphos_reason <- "low"
} else{
  tphos_score <- 2
  tphos_reason <- "moderate"
}


chla <- trophic_df$LongtermAvg[trophic_df$FULL_CHARACTERISTIC_NAME == "CHLOROPHYLL A OW T"]

if(is.na(chla)){
  chla_score <- "0"
  chla_reason <- "unsampled"
} else if(chla > 8){
  chla_score <- 1
  chla_reason <- "high"
} else if(chla < 2){
  chla_score <- 3
  chla_reason <- "low"
} else{
  chla_score <- 2
  chla_reason <- "moderate"
}


trophic_sum <- secchi_score + tphos_score + chla_score


if(trophic_sum == 3){
  overall_trophic_state <- "eutrophic"
  ts_descriptor <- "highly productive"
} else if(trophic_sum == 9){
  overall_trophic_state <- "oligotrophic"
  ts_descriptor <- "moderately to highly productive"
} else if(trophic_sum == 6){
  overall_trophic_state <- "mesotrophic"
  ts_descriptor <- "moderately productive"
} else if(trophic_sum %in% c(4,5)){
  overall_trophic_state <- "mesoeutrophic"
  ts_descriptor <- "moderately unproductive"
} else if(trophic_sum %in% c(7,8)){
  overall_trophic_state <- "mesoligotrophic"
  ts_descriptor <- "unproductive"
}

###soluble data

soluble <- lake1 %>% 
    distinct(SAMPLE_ID, INFO_TYPE, Characteristic.Name, Result.Sample.Fraction, Result.Value, .keep_all = TRUE) %>% 
  filter(year == report_year)

thisyear_obs <- nrow(soluble)

soluble$Result.Sample.Fraction <- replace_na(soluble$Result.Sample.Fraction, "None")

soluble <- soluble %>% 
  mutate(FULL_CHARACTERISTIC_NAME = paste(Characteristic.Name, INFO_TYPE, Result.Sample.Fraction, sep=" ")) %>%
  filter(FULL_CHARACTERISTIC_NAME %in% c("PHOSPHORUS OW T",
                                         "PHOSPHORUS OW D",
                                         "NITROGEN, TOTAL OW T",
                                         "NITROGEN, TOTAL DISSOLVED OW D")) %>% 
  select(SAMPLE_ID, FULL_CHARACTERISTIC_NAME, Result.Value) %>% 
  group_by(SAMPLE_ID, FULL_CHARACTERISTIC_NAME) %>% 
  summarize(Result.Value = mean(as.numeric(Result.Value))) %>% 
  ungroup()

soluble$FULL_CHARACTERISTIC_NAME[soluble$FULL_CHARACTERISTIC_NAME == "PHOSPHORUS OW T"] <- "TP"
soluble$FULL_CHARACTERISTIC_NAME[soluble$FULL_CHARACTERISTIC_NAME == "PHOSPHORUS OW D"] <- "TDP"
soluble$FULL_CHARACTERISTIC_NAME[soluble$FULL_CHARACTERISTIC_NAME == "NITROGEN, TOTAL OW T"] <- "TN"
soluble$FULL_CHARACTERISTIC_NAME[soluble$FULL_CHARACTERISTIC_NAME == "NITROGEN, TOTAL DISSOLVED OW D"] <- "TDN"

soluble_df <-  soluble %>% 
  spread(FULL_CHARACTERISTIC_NAME, Result.Value) %>% 
  mutate(TDN_TN = TDN/TN) %>% 
  mutate(TDP_TP = TDP/TP)

max_p_ratio <- max(soluble_df$TDP_TP)
avg_p_ratio <- mean(soluble_df$TDP_TP)
max_n_ratio <- max(soluble_df$TDN_TN)
avg_n_ratio <- mean(soluble_df$TDN_TN)

if(thisyear_obs == 0){
  soluble_nutrients <- paste("Soluble nutrients were not analyzed in ", report_year, ".", sep="")
} else{
  soluble_nutrients <- paste("Soluble nutrients were analyzed in ", report_year, ".", sep="")
}

if(avg_p_ratio > 0.5){
  soluble_p <- "Most of the phosphorus in the lake is soluble, indicating a high potential for more algae growth."
} else if(max_p_ratio > 0.5){
  soluble_p <- "Some of the phosphorus in the lake is soluble, indicating some potential for more algae growth."
} else if(max_p_ratio > 0.25){
  soluble_p <- "Most of the phosphorus in the lake is not soluble, indicating a low potential for additional algae growth."
} else{
  soluble_p <- "Most of the phosphorus in the lake is not soluble, suggesting most of the phosphorus is bound in algae cells."
}

if(avg_n_ratio > 0.5){
  soluble_n <- "Most of the nitrogen in the lake is soluble."
} else if(max_n_ratio > 0.5){
  soluble_n <- "Some of the nitrogen in the lake is soluble."
} else if(max_n_ratio > 0.25){
  soluble_n <- "Most of the nitrogen in the lake is not soluble."
} else{
  soluble_n <- "Most of the nitrogen in the lake is not soluble."
}


#ph, conductivity, color, nitrogen

water_desc <- lake1 %>% 
    distinct(SAMPLE_ID, INFO_TYPE, Characteristic.Name, Result.Sample.Fraction, Result.Value, .keep_all = TRUE) %>% 
  filter(year == report_year)

water_desc$Result.Sample.Fraction <- replace_na(water_desc$Result.Sample.Fraction, "None")

water_desc <- water_desc %>% 
  mutate(FULL_CHARACTERISTIC_NAME = paste(Characteristic.Name, INFO_TYPE, Result.Sample.Fraction, sep=" ")) %>%
  filter(FULL_CHARACTERISTIC_NAME %in% c("DEPTH, SECCHI DISK DEPTH SD None",
                                         "PHOSPHORUS OW T",
                                         "CHLOROPHYLL A OW T",
                                         "PH OW None",
                                          "TRUE COLOR OW T",
                                         "NITROGEN, TOTAL OW T",
                                         "SPECIFIC CONDUCTANCE OW None",
                                         "QA OW None",
                                         "QB OW None",
                                         "QC OW None",
                                         "CHLORIDE OW T",
                                         "TEMPERATURE, WATER OW None",
                                         "TEMPERATURE, WATER BS None"
                                         )) %>% 
  select(FULL_CHARACTERISTIC_NAME, Result.Value) %>% 
  group_by(FULL_CHARACTERISTIC_NAME) %>% 
  summarize(ThisYearAvg = mean(as.numeric(Result.Value)), 
            standard_dev = sd(as.numeric(Result.Value))) %>% 
  ungroup()


ph_avg <- water_desc$ThisYearAvg[water_desc$FULL_CHARACTERISTIC_NAME == "PH OW None"]
cond_avg <- water_desc$ThisYearAvg[water_desc$FULL_CHARACTERISTIC_NAME == "SPECIFIC CONDUCTANCE OW None"]
color_avg <- water_desc$ThisYearAvg[water_desc$FULL_CHARACTERISTIC_NAME == "TRUE COLOR OW T"]
tnit_avg <- water_desc$ThisYearAvg[water_desc$FULL_CHARACTERISTIC_NAME == "NITROGEN, TOTAL OW T"]

if(ph_avg < 6){
  ph_descriptor <- "acidic"
} else if(ph_avg < 6.5){
  ph_descriptor <- "slightly acidic"
} else if(ph_avg < 7.5){
  ph_descriptor <- "a near neutral pH"
} else if(ph_avg < 8){
  ph_descriptor <- "slightly alkaline or basic"
} else{
  ph_descriptor <- "highly alkaline or basic"
}

if(cond_avg < 100){
  cond_descriptor <- "soft water"
} else if(cond_avg < 250){
  cond_descriptor <- "intermediate hardness water"
} else if(cond_avg < 400){
  cond_descriptor <- "hard water"
} else{
  cond_descriptor <- "extremely hard water"
}

if(color_avg < 10){
  color_descriptor <- "low water color"
} else if(color_avg < 30){
  color_descriptor <- "moderately low water color"
} else if(color_avg < 50){
  color_descriptor <- "moderately high water color"
} else{
  color_descriptor <- "high water color"
}

if(tnit_avg < 0.3){
  tnit_descriptor <- "low nitrogen levels"
} else if(tnit_avg < 0.5){
  tnit_descriptor <- "moderately low nitrogen levels"
} else if(tnit_avg < 0.8){
  tnit_descriptor <- "moderately high nitrogen levels"
} else{
  tnit_descriptor <- "high nitrogen levels"
}


```


**Q. What is the condition of the lake?**

A. `r titles[1]` continues to be  `r overall_trophic_state`, or `r ts_descriptor`, based on `r secchi_reason` water clarity, `r chla_reason` algae levels (chlorophyll a), and `r tphos_reason` nutrient (phosphorous) levels. `r soluble_nutrients` `r soluble_p` `r soluble_n` The waterbody is `r ph_descriptor`, with `r cond_descriptor`, `r color_descriptor`, and `r tnit_descriptor`. 

```{r prevyears_summary, echo=FALSE, message=FALSE, warning=FALSE}
prev_decade <- lake1 %>% 
    distinct(SAMPLE_ID, INFO_TYPE, Characteristic.Name, Result.Sample.Fraction, Result.Value, .keep_all = TRUE) %>% 
  filter(as.numeric(year) < as.numeric(report_year)) %>% 
  filter(as.numeric(year) >= (as.numeric(report_year) - 10))

prev_decade$Result.Sample.Fraction <- replace_na(prev_decade$Result.Sample.Fraction, "None")

prev_decade <- prev_decade %>% 
  mutate(FULL_CHARACTERISTIC_NAME = paste(Characteristic.Name, INFO_TYPE, Result.Sample.Fraction, sep=" ")) %>%
  filter(FULL_CHARACTERISTIC_NAME %in% c("DEPTH, SECCHI DISK DEPTH SD None",
                                         "PHOSPHORUS OW T",
                                         "CHLOROPHYLL A OW T",
                                         "PH OW None",
                                          "TRUE COLOR OW T",
                                         "NITROGEN, TOTAL OW T",
                                         "SPECIFIC CONDUCTANCE OW None",
                                         "QA OW None",
                                         "QB OW None",
                                         "QC OW None",
                                         "CHLORIDE OW T",
                                         "TEMPERATURE, WATER OW None",
                                         "TEMPERATURE, WATER BS None"
                                         )) %>% 
  select(FULL_CHARACTERISTIC_NAME, Result.Value) %>% 
  filter(Result.Value != "NA") %>% 
  group_by(FULL_CHARACTERISTIC_NAME) %>% 
  summarize(TenYearAvg = mean(as.numeric(Result.Value))) %>% 
  ungroup()

trends <- full_join(prev_decade, water_desc) %>% 
  mutate(change = abs(TenYearAvg - ThisYearAvg)) %>% 
  mutate(sig = change - standard_dev)

FULL_CHARACTERISTIC_NAME <- c("DEPTH, SECCHI DISK DEPTH SD None",
                                         "PHOSPHORUS OW T",
                                         "CHLOROPHYLL A OW T",
                                         "PH OW None",
                                          "TRUE COLOR OW T",
                                         "NITROGEN, TOTAL OW T",
                                         "SPECIFIC CONDUCTANCE OW None",
                                         "QA OW None",
                                         "QB OW None",
                                         "QC OW None",
                                         "CHLORIDE OW T",
                                         "TEMPERATURE, WATER OW None",
                                         "TEMPERATURE, WATER BS None"
                                         )

ordering <- data.frame(FULL_CHARACTERISTIC_NAME)

trends <- left_join(ordering, trends)


for(i in 1:nrow(trends)){
  if(is.na(trends$sig[i])){
  trends$comparison[i] <- "insufficient data"
  } else if(trends$sig[i] < 0){
    trends$comparison[i] <- "similar"
  } else if(trends$sig[i] > 0){
    if(trends$TenYearAvg[i] > trends$ThisYearAvg[i]){
       trends$comparison[i] <- "lower"
    } else if (trends$TenYearAvg[i] < trends$ThisYearAvg[i]){
      trends$comparison[i] <- "higher"
    }
  }
}




trends_q <- trends %>% 
  filter(FULL_CHARACTERISTIC_NAME %in% c("QA OW None", "QB OW None", "QC OW None"))

trends <- trends %>% 
  filter(!FULL_CHARACTERISTIC_NAME %in% c("QA OW None", "QB OW None", "QC OW None"))

trends_q$FULL_CHARACTERISTIC_NAME[trends_q$FULL_CHARACTERISTIC_NAME == "QA OW None"] <- "water quality evaluation"
trends_q$FULL_CHARACTERISTIC_NAME[trends_q$FULL_CHARACTERISTIC_NAME == "QB OW None"] <- "aquatic plant coverage"
trends_q$FULL_CHARACTERISTIC_NAME[trends_q$FULL_CHARACTERISTIC_NAME == "QC OW None"] <- "recreational evaluation"


trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "DEPTH, SECCHI DISK DEPTH SD None"] <- "water clarity (secchi)"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "PHOSPHORUS OW T"] <- "total phosphorous"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "CHLOROPHYLL A OW T"] <- "extracted chlorophyll a"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "PH OW None"] <- "pH"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME ==  "TRUE COLOR OW T"] <- "color"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME ==  "NITROGEN, TOTAL OW T"] <- "total nitrogen"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME ==  "SPECIFIC CONDUCTANCE OW None"] <- "conductivity"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "CHLORIDE OW T"] <- "chloride"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "TEMPERATURE, WATER OW None"] <- "surface water temperature"
trends$FULL_CHARACTERISTIC_NAME[trends$FULL_CHARACTERISTIC_NAME == "TEMPERATURE, WATER BS None"] <- "deep water temperature"


###make this into for loop eventually

insuff <- trends %>% 
  filter(comparison == "insufficient data") 
if(nrow(insuff) > 1){
  heads <- toString(head(insuff$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(insuff$FULL_CHARACTERISTIC_NAME, 1))
  insuff_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(insuff) == 1){
  insuff_list <- insuff$FULL_CHARACTERISTIC_NAME[1]
} 

insuff_q <- trends_q %>% 
  filter(comparison == "insufficient data")
if(nrow(insuff_q) > 1){
  heads <- toString(head(insuff_q$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(insuff_q$FULL_CHARACTERISTIC_NAME, 1))
  insuff_q_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(insuff) == 1){
  insuff__q_list <- insuff_q$FULL_CHARACTERISTIC_NAME[1]
} 

sim <- trends %>% 
  filter(comparison == "similar")
if(nrow(sim) > 1){
  heads <- toString(head(sim$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(sim$FULL_CHARACTERISTIC_NAME, 1))
  sim_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(sim) == 1){
  sim_list <- sim$FULL_CHARACTERISTIC_NAME[1]
} 

sim_q <- trends_q %>% 
  filter(comparison == "similar")
if(nrow(sim_q) > 1){
  heads <- toString(head(sim_q$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(sim_q$FULL_CHARACTERISTIC_NAME, 1))
  sim_q_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(sim_q) == 1){
  sim_q_list <- sim_q$FULL_CHARACTERISTIC_NAME[1]
} 

low <- trends %>% 
  filter(comparison == "lower")
if(nrow(low) > 1){
  heads <- toString(head(low$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(low$FULL_CHARACTERISTIC_NAME, 1))
  low_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(low) == 1){
  low_list <- low$FULL_CHARACTERISTIC_NAME[1]
} 

low_q <- trends_q %>% 
  filter(comparison == "lower")
if(nrow(low_q) > 1){
  heads <- toString(head(low_q$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(low_q$FULL_CHARACTERISTIC_NAME, 1))
  low_q_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(low_q) == 1){
  low_q_list <- low_q$FULL_CHARACTERISTIC_NAME[1]
} 


high <- trends %>% 
  filter(comparison == "higher")
if(nrow(high) > 1){
  heads <- toString(head(high$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(high$FULL_CHARACTERISTIC_NAME, 1))
  high_list <- paste(heads, "and", tails, sep=" ")
}else if(nrow(high) == 1){
  high_list <- high$FULL_CHARACTERISTIC_NAME[1]
} 

high_q <- trends_q %>% 
  filter(comparison == "higher")
if(nrow(high_q) > 1){
  heads <- toString(head(high_q$FULL_CHARACTERISTIC_NAME, -1))
  tails <- toString(tail(high_q$FULL_CHARACTERISTIC_NAME, 1))
  high_q_list <- paste(heads, "and", tails, sep=" ")
} else if(nrow(high_q) == 1){
  high_q_list <- high_q$FULL_CHARACTERISTIC_NAME[1]
} 


if(nrow(high) == 0){
    if(nrow(high_q) ==0){
      higher_sentence <- ""
    } else if(nrow(high_q) == 1){
      higher_sentence <- paste0("Compared to previous years, ", high_q_list, " was less favorable in ",         report_year, ".") 
    } else if(nrow(high_q) > 1){
      higher_sentence <- paste0("Compared to previous years, ", high_q_list, " were less favorable in ",        report_year, ".")}
} else if(nrow(high) == 1){
    if(nrow(high_q) == 0){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " was higher in ",           
      report_year, ".")
    } else if(nrow(high_q) == 1){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " was higher in ",        
      report_year, " and ", high_q_list, " was less favorable in ", report_year, ".")
    } else if(nrow(high_q) > 1){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " was higher in ", 
      report_year, " and ", high_q_list, " was less favorable in ", report_year, ".")}
} else if(nrow(high) > 1){
    if(nrow(high_q) == 0){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " were higher in ",        
      report_year, ".")
    } else if(nrow(high_q) == 1){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " were higher in ",        
      report_year, " and ", high_q_list, " was less favorable in ", report_year, ".")
    } else if(nrow(high_q) > 1){
      higher_sentence <- paste0("Compared to previous years, ", high_list, " were higher in ", 
      report_year, " and ", high_q_list, " were less favorable in ", report_year, ".")}
}


if(nrow(low) == 0){
    if(nrow(low_q) == 0){
      lower_sentence <- " "
    } else if(nrow(low_q) == 1){
      lower_sentence <- paste0("Compared to previous years, ", low_q_list, " was more favorable in ",         report_year, ".") 
    } else if(nrow(low_q) > 1){
      lower_sentence <- paste0("Compared to previous years, ", low_q_list, " were more favorable in ",        report_year, ".")}
} else if(nrow(low) == 1){
    if(nrow(low_q) == 0){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " was lower in ",           
      report_year, ".")
    } else if(nrow(low_q) == 1){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " was lower in ",        
      report_year, " and ", low_q_list, " was more favorable in ", report_year, ".")
    } else if(nrow(low_q) > 1){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " was lower in ", 
      report_year, " and ", low_q_list, " was more favorable in ", report_year, ".")}
} else if(nrow(low) > 1){
    if(nrow(low_q) == 0){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " were lower in ",        
      report_year, ".")
    } else if(nrow(low_q) == 1){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " were lower in ",        
      report_year, " and ", low_q_list, " was more favorable in ", report_year, ".")
    } else if(nrow(low_q) > 1){
      lower_sentence <- paste0("Compared to previous years, ", low_list, " were lower in ", 
      report_year, " and ", low_q_list, " was more favorable in ", report_year, ".")}
}


```
**Q. How did this year compare to previous years?**

A. `r higher_sentence` `r lower_sentence`

**Q. How does this lake compare to other New York lakes?**

A. 

**Q. Are there any (statistically significant) trends?**

A.

**Q. Has the lake experienced harmful algal blooms (HABs)?**

A.

```{r invasive_summary, echo=FALSE, message=FALSE, warning=FALSE}
lake_info <- read.csv(file = "C:/Users/amtweitm/OneDrive - New York State Office of Information Technology Services/Documents/R/CSLAP_reporting/data/lake_info.csv") %>% 
  filter(LakeID == ids)

if(lake_info$Invasive.Animals.[1] %in% c(NA, "none reported", "")){
  inv_animals <- 0
} else{inv_animals <- 1}

if(inv_animals == 1){
 lake_info$Invasive.Animals.[1] <- str_replace_all(lake_info$Invasive.Animals.[1], ";", ",")
}

if(lake_info$Invasive.Plants.[1] %in% c(NA, "none reported", "")){
  inv_plants <- 0
} else{inv_plants <- 1}

if(inv_plants == 1){
  lake_info$Invasive.Plants.[1] <- str_replace_all(lake_info$Invasive.Plants.[1], ";", ",")
}

invasives <- inv_animals + inv_plants

if(invasives == 0){
  invasives_present <- "No invasive species have been reported in this waterbody. "
} else if(invasives > 0){
    if(inv_animals == 0){
      invasives_present <- paste("Invasive species have been reported in this waterbody. Aquatic invasive plant and/or animal species reported include: ", lake_info$Invasive.Plants.[1], ".", sep="")
    } else if (inv_plants == 0){
      invasives_present <- paste("Invasive species have been reported in this waterbody. Aquatic invasive plant and/or animal species reported include: ", lake_info$Invasive.Animals.[1],  ".", sep="")
    } else{
      invasives_present <- paste("Invasive species have been reported in this waterbody. Aquatic invasive plant and/or animal species reported include: ", lake_info$Invasive.Animals.[1], ", ", lake_info$Invasive.Plants.[1], ".", sep="")
    }
}

calcium <- lake1 %>% 
  filter(Characteristic.Name == "CALCIUM") %>% 
  filter(!is.na(Result.Value)) %>% 
  group_by(Characteristic.Name, year) %>% 
  summarize(avg_calc = mean(as.numeric(Result.Value))) %>% 
  ungroup() %>% 
  group_by(Characteristic.Name) %>% 
  summarize(Longterm_Avg = mean(avg_calc))

  

if(invasives > 0){
 inv_vulnerability <- "This waterbody has high vulnerability for new invasives, based on invasive species already introduced and calcium levels." 
} else if(is.na(calcium$Longterm_Avg[1])){
  inv_vulnerability <- "This waterbody has unknown vulnerability."
} else if(calcium$Longterm_Avg[1] > 30){
  inv_vulnerability <- "This waterbody has high vulnerability for new invasives, based on calcium levels."
} else if(calcium$Longterm_Avg[1] >= 15 & calcium$Longterm_Avg[1] <= 30){
  inv_vulnerability <- "This waterbody has moderate vulnerability, based on calcium levels."
} else if(calcium$Longterm_Avg[1] < 15){
  inv_vulnerability <- "This waterbody has low vulnerability, based on calcium levels."
}

```

**Q. Have any aquatic invasive species (AIS) been reported?**

A. `r invasives_present` `r inv_vulnerability`
